Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-12-12 Manuscript/Imputation_of_Incomplete_Multilevel_Data.tex
tables

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

@@

\author{
Hanne Oberman\\Utrecht University \And Johanna Munoz Avila\\University
Medical Center Utrecht \AND Valentijn de Jong\\University Medical
Center Utrecht \And Gerko Vink\\Utrecht University \AND Thomas
Debray\\University Medical Center Utrecht
}
\title{Imputation of Incomplete Multilevel Data with \pkg{mice}}

\Plainauthor{Hanne Oberman, Johanna Munoz Avila, Valentijn de
Jong, Gerko Vink, Thomas Debray}
\Plaintitle{Imputation of Incomplete Multilevel Data with mice}
\Shorttitle{Multilevel \pkg{mice}}

@@

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

\begin{table}[tb]
\caption{Concepts in multilevel methods}
\label{tab:clus}
\centering
\begin{tabular}{ll}
\hline
\textbf{Concept} & \textbf{Details}   \\
\hline
ICC                 & The variability due to clustering is often measured by means of the \\
                    & intraclass coefficient (ICC). The ICC can be seen as the percentage \\
                    & of variance that can be attributed to the cluster-level, where a high \\
                    & ICC would indicate that a lot of variability is due to the cluster \\
                    & structure. \\
Random effect       & Multilevel models typically accommodate for variability by including \\
                    & a separate group mean for each cluster. In addition to random \\
                    & intercepts, multilevel models can also include random coefficients \\
                    & and heterogeneous residual error variances across clusters [see e.g. \\
                    & @gelm06, @hox17 and @jong21]. [TODO: add stratification.] \\
\hline
\end{tabular}
\end{table}

\hypertarget{missingness-in-multilevel-data}{%
\subsection{Missingness in multilevel
data}\label{missingness-in-multilevel-data}}

Like any other dataset, clustered datasets are prone to missing data.
Several strategies can be used to handle missing data, including
complete case analysis and imputation. We focus on the latter approach
and discuss statistical methods for replacing the missing data with one
or more plausible values. Afterwards, the completed data can be analyzed
as if they were completely observed. In contrast to single imputation
(where missing data are only replaced once), multiple imputation allows
to preserve uncertainty due to missingness and is therefore recommended
(c.f. Rubin 1976).

When clustered datasets are affected by missing values, we can
distinguish between two types of missing data: sporadic missingness and
systematic missingness \citep{resc13}. Sporadic missingness arises when
variables are missing for some but not all of the units in a cluster
\citep{buur18, jola18}. For example, it is possible that test results
are missing for several students in one or more classes. {[}TODO:
Provide an example for one of the case studies below.{]} When all
observations are missing within one or more clusters, data are
systematically missing. {[}TODO: Refer to Figure 1 and put
interpretation in the figure caption.{]}

\begin{CodeChunk}
\begin{figure}

@@

\caption[Missingness in multilevel data]{Missingness in multilevel data}\label{fig:patterns}
\end{figure}
\end{CodeChunk}

@@

\begin{table}[tb]
\caption{Notation}
\label{tab:mod}
\centering
\begin{tabular}{ll}
\hline
\textbf{Concept} & \textbf{Details}   \\
\hline
& [TODO: explain \pkg{lme4} notation here] \\
\hline
\end{tabular}
\end{table}

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     4.932     0.274    18.012     4.420     0.000    19.514     0.964 
sex             0.859     0.284     3.026     4.216     0.036    37.601     0.981

                            Estimate 
Intercept~~Intercept|school    0.265 
Residual~~Residual             1.007 
ICC|school                     0.213

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     5.054     0.266    19.029     4.554     0.000    14.936     0.954 
sex             0.732     0.306     2.393     4.233     0.071    34.886     0.980

                            Estimate 
Intercept~~Intercept|school    0.318 
Residual~~Residual             1.257 
ICC|school                     0.208

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     4.714     0.193    24.432     5.032     0.000     8.224     0.919 
sex             1.286     0.288     4.460     4.218     0.010    37.215     0.981

                            Estimate 
Intercept~~Intercept|school    0.299 
Residual~~Residual             1.050 
ICC|school                     0.228

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- glmer(mod, family = "binomial", data = na.omit(dat)) # fit the model
R> tidy(fit, conf.int = TRUE, exponentiate = TRUE)             # print estimates
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 11 x 9
   effect   group term        estim~1 std.er~2 stati~3   p.value conf.~4 conf.~5
   <chr>    <chr> <chr>         <dbl>    <dbl>   <dbl>     <dbl>   <dbl>   <dbl>
 1 fixed    <NA>  (Intercept)  0.0863  0.0182   -11.6   2.99e-31  0.0571   0.130
 2 fixed    <NA>  typeRCT      0.757   0.137     -1.54  1.22e- 1  0.531    1.08 
 3 fixed    <NA>  age          1.03    0.00265   12.9   7.40e-38  1.03     1.04 
 4 fixed    <NA>  motor_scor~  0.651   0.0732    -3.82  1.34e- 4  0.522    0.811
 5 fixed    <NA>  motor_scor~  0.489   0.0555    -6.30  2.97e-10  0.391    0.611
 6 fixed    <NA>  motor_scor~  0.274   0.0321   -11.0   2.28e-28  0.218    0.345
 7 fixed    <NA>  pupilNone    3.20    0.317     11.7   8.18e-32  2.63     3.88 
 8 fixed    <NA>  pupilOne     1.75    0.195      5.06  4.27e- 7  1.41     2.18 
 9 fixed    <NA>  ctIII        2.41    0.268      7.89  3.05e-15  1.94     2.99 
10 fixed    <NA>  ctIV/V       2.30    0.214      8.95  3.55e-19  1.92     2.76 
11 ran_pars name  sd__(Inter~  0.230  NA         NA    NA        NA       NA    
# ... with abbreviated variable names 1: estimate, 2: std.error, 3: statistic,
#   4: conf.low, 5: conf.high
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- imp %>% 
+   with(glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial")) 
R> tidy(pool(fit))
\end{CodeInput}
\begin{CodeOutput}
                     term    estimate   std.error  statistic      p.value
1             (Intercept) -2.33813790 0.340459131  -6.867602 6.978862e-12
2                    type -0.41509929 0.179936773  -2.306918 2.107792e-02
3                     age  0.03034813 0.001561834  19.431086 0.000000e+00
4 as.factor(motor_score)2 -0.66052409 0.068846376  -9.594174 0.000000e+00
5 as.factor(motor_score)3 -1.05030912 0.070059114 -14.991756 0.000000e+00
6 as.factor(motor_score)4 -1.49637069 0.071492961 -20.930322 0.000000e+00
7                   pupil  0.48922641 0.038536483  12.695149 0.000000e+00
8                      ct  0.42634415 0.028388807  15.018037 0.000000e+00
             b          df dfcom          fmi       lambda m          riv
1 9.395965e-04  8668.10341 11013 0.0099557149 0.0097273074 5 0.0098228573
2 1.729929e-05 10991.51028 11013 0.0008229571 0.0006411648 5 0.0006415761
3 8.332782e-09 10482.95010 11013 0.0042891736 0.0040992241 5 0.0041160969
4 5.649577e-05  6979.24803 11013 0.0145856016 0.0143032580 5 0.0145108098
5 3.003051e-05  9526.87220 11013 0.0075503299 0.0073420043 5 0.0073963080
6 3.329522e-05  9362.42789 11013 0.0080288174 0.0078169354 5 0.0078785213
7 3.337070e-04    54.63795 11013 0.2949938851 0.2696512493 5 0.3692088868
8 1.180517e-05  5893.23780 11013 0.0179108151 0.0175775781 5 0.0178920775
          ubar
1 1.147849e-01
2 3.235648e-02
3 2.429325e-06
4 4.672029e-03
5 4.872243e-03
6 5.071289e-03
7 1.084612e-03
8 7.917581e-04
\end{CodeOutput}
\begin{CodeInput}
R> as.mitml.result(fit)
\end{CodeInput}
\begin{CodeOutput}
[[1]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
      AIC       BIC    logLik  deviance  df.resid 
10522.944 10588.713 -5252.472 10504.944     11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2913  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.32713                 -0.41953                  0.03031  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.65651                 -1.04719                 -1.49327  
                  pupil                       ct  
                0.48990                  0.42335  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[2]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
      AIC       BIC    logLik  deviance  df.resid 
10525.723 10591.492 -5253.861 10507.723     11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2937  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.30150                 -0.41940                  0.03024  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.65497                 -1.04451                 -1.49454  
                  pupil                       ct  
                0.47281                  0.42888  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[3]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
     AIC      BIC   logLik deviance df.resid 
10512.28 10578.05 -5247.14 10494.28    11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2852  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.35320                 -0.41028                  0.03034  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.66933                 -1.05474                 -1.50379  
                  pupil                       ct  
                0.49491                  0.42442  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[4]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
     AIC      BIC   logLik deviance df.resid 
10530.02 10595.79 -5256.01 10512.02    11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2923  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.32670                 -0.41286                  0.03037  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.66801                 -1.04770                 -1.50070  
                  pupil                       ct  
                0.47210                  0.43109  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[5]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
      AIC       BIC    logLik  deviance  df.resid 
10488.615 10554.383 -5235.307 10470.615     11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2939  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.38215                 -0.41343                  0.03049  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.65379                 -1.05740                 -1.48956  
                  pupil                       ct  
                0.51641                  0.42397  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

@@

Impute the data:

\begin{CodeChunk}
\begin{CodeInput}
R> imp <- mice( hivdata, # dataset with missing values
+                         m = 10,   # number of imputations
+                         maxit = 1,
+                         seed = 1234, #seed attached to the dataID
+                         meth = meth, #imputation method vector
+                         pred = pred, #imputation predictors matrix
+                         print = T,
+                         meta_method="reml",
+                         pmm=FALSE)
\end{CodeInput}
\begin{CodeOutput}

 iter imp variable
  1   1  hiv
  1   2  hiv
  1   3  hiv
  1   4  hiv
  1   5  hiv
  1   6  hiv
  1   7  hiv
  1   8  hiv
  1   9  hiv
  1   10  hiv
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{itemize}
\item
  Adding some kind of help function to mice that suggests a suitable
  predictor matrix to the user, given a certain analysis model.
\item
  Adding a \texttt{multilevel\_ampute()} wrapper function in mice.
\item
  Exporting \texttt{mids} objects to other packages like \texttt{lme4}
  or \texttt{coxme}?
\item
  Adding a ICC=0 dataset to show that even if there is no clustering it
  doesn't hurt.
\item
  Show use case for deductive imputation for cluster level variables?
\item
  env dump in repo
\end{itemize}

>>>>

@@

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Hanne Oberman\\Utrecht University \And Johanna Munoz\\University Medical
Center Utrecht \AND Valentijn de Jong\\University Medical Center
Utrecht \And Gerko Vink\\Utrecht University \AND Thomas
Debray\\University Medical Center Utrecht
}
\title{Imputation of Incomplete Multilevel Data with \pkg{mice}}

\Plainauthor{Hanne Oberman, Johanna Munoz, Valentijn de Jong, Gerko
Vink, Thomas Debray}
\Plaintitle{Imputation of Incomplete Multilevel Data with mice}
\Shorttitle{Multilevel \pkg{mice}}

@@

% Pandoc syntax highlighting

% Pandoc citation processing

@@

\begin{table}[tb]
\caption{Concepts in multilevel methods}
\label{tab:clus}
\centering
\begin{tabular}{ll}
\hline
\textbf{Concept} & \textbf{Details}   \\
\hline
Sample unit         & Units of the population from which measurements are taken in a sample.\\                           & These should be distinct and not overlapping, e.g., patients, students.\\ 
Hierarchical levels & Data are grouped into clusters at different levels. A three-level \\                               & hierarchical model could occur by collecting observations of math scores\\                         & (lower level) which can be measured over the whole school year for each\\                          & student, and then grouped by student (middle level) and these in turn can\\                        & be aggregated into classrooms (high level).\\
Fixed effect        & Here we assume that the values of an independent variable are fixed, i.e., \\ 
  & the values observed in the study are representative of all values in the \\                        & population we are interested in. In the case of a fixed factor variable, for \\                     & example treatment type, we are only concerned with assessing the difference \\ 
                & in the dependent variable y e.g., blood pressure between treatments A and B.\\                     & There may be additional treatments, but based on the objective of the study,\\                     & which in this case is the equivalence effect of A and B, the other treatment\\                     & types are irrelevant and therefore the treatment type is included as a fixed \\                     & effect with only the A and B levels in the hierarchical model. \\
Random effect       & The values of an independent variable are assumed to be randomly drawn from \\                     & a population with more values. In a random factor, the levels observed in the\\                     & study represent a sample of all possible levels, e.g., hospitals, classrooms.\\                     & For example, in a study to predict the blood pressure of patients on \\  
& admission we might select only certain hospitals that are representative of \\                     & the entire population of hospitals. Here we are not interested in measuring \\ 
& the difference of y between individual hospitals, but rather the variation of\\                     & y between hospitals, to create a model that can be generalized to the entire\\                     & population of hospitals. \\

ICC                 & The variability due to clustering is often measured by means of the \\
                    & intraclass coefficient (ICC). The ICC can be seen as the percentage \\
                    & of variance that can be attributed to the cluster-level, where a high \\
                    & ICC would indicate that a lot of variability is due to the cluster \\
                    & structure. \\
Random effect       & Multilevel models typically accommodate for variability by including \\
                    & a separate group mean for each cluster. In addition to random \\
                    & intercepts, multilevel models can also include random coefficients \\
                    & and heterogeneous residual error variances across clusters [see e.g. \\
                    & @gelm06, @hox17 and @jong21]. [TODO: add stratification.] \\
\hline
\end{tabular}
\end{table}

\begin{verbatim}

## Missingness in multilevel data

<!-- Maybe besides CC and imputation name also likelihood based methods -->
Like any other dataset, clustered datasets are prone to missing data. Several strategies can be used to handle missing data, including complete case analysis and imputation. We focus on the latter approach and discuss statistical methods for replacing the missing data with one or more plausible values. Afterwards, the completed data can be analyzed as if they were completely observed. In contrast to single imputation (where missing data are only replaced once), multiple imputation allows to preserve uncertainty due to missingness and is therefore recommended (c.f. Rubin 1976).

When clustered datasets are affected by missing values, we can distinguish between two types of missing data: sporadic missingness and systematic missingness [@resc13]. Sporadic missingness arises when variables are missing for some but not all of the units in a cluster [@buur18; @jola18]. For example, it is possible that test results are missing for several students in one or more classes. [TODO: Provide an example for one of the case studies below.] When all observations are missing within one or more clusters, data are systematically missing. [TODO: Refer to Figure 1 and put interpretation in the figure caption.]

<!-- Systematic missingness implies that one or more variables are never observed in a certain cluster. With sporadic missingness there may be observed data for some but not all units in a cluster [@buur18; @jola18]. We have visualized this difference in Figure 1, which shows an $n \times p$ set $\mathbf{X} = X_1, \dots, X_p$, with $n$ units distributed over $N$ clusters and $p$ variables.  -->

```{=latex}
\begin{CodeChunk}
\begin{figure}

@@

\caption[Missingness in multilevel data]{Missingness in multilevel data}\label{fig:patterns}
\end{figure}
\end{CodeChunk}
\end{verbatim}

@@

\begin{table}[tb]
\caption{Notation}
\label{tab:mod}
\centering
\begin{tabular}{lll}
\hline
\textbf{Formula lme4} & \textbf{Details}   \\
\hline
y$\sim$ x1 + (1 | g1)                                       & Fixed x1 predictor with random intercept  \\
                                                         & varying among g\\
y$\sim$x1*x2+ (1| g1)                                      & Interactions of x1 and x2 only in fixed effect\\
y$\sim$x1*x2+ (x2| g1)                                     & Interactions of x1 and x2 only in fixed effect\\
                                                         & with slope of x2 randomly varying among g1\\
y$\sim$x1*x2+ (x1*x2| g1)                                  & variance-covariance matrix estimated only with \\
                                                         & the variance terms of intercept, slope of x1, \\
                                                         & slope of x2 and interaction x1*x2\\
y$\sim$x1*x2+ (x1| g1)+ (x2| g1)                           & variance-covariance matrix estimated separately,\\
                                                         & i.e, one for intercept and x1 and another for \\
                                                         & intercept and x2\\
y$\sim$x1 + (x1 | g1) or 1 + x1 + (1 + x1 | g1)            & Fixed x1 with correlated random intercept and \\
                                                         & random slope of x\\
y$\sim$x1 + (x1 || g1) or 1 + x1 + (1 | g1) + (0 + x1 | g1) & Fixed x1 with uncorrelated random intercept\\
                                                         & and random slope of x1\\  
y$\sim$(1 | g1) + (1 | g2)                                 & Random intercept varying among g1 and among g2\\
y$\sim$(1 | g1/g2) or ~(1|g1)+(1|g1:g2)                    & Random intercept varying among g1 and g2 
                                                         & withing g1\\

& [TODO: explain \pkg{lme4} notation here
https://datalorax.github.io/equatiomatic/articles/lmer4-lmer.html
https://cran.r-project.org/web/packages/lme4/vignettes/lmer.pdf
] \\
\hline
\end{tabular}
\end{table}

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     5.012     0.295    16.994     4.362     0.000    22.587     0.969 
sex             0.695     0.251     2.768     4.287     0.047    28.390     0.975

                            Estimate 
Intercept~~Intercept|school    0.266 
Residual~~Residual             1.035 
ICC|school                     0.208

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     4.915     0.217    22.642     4.926     0.000     9.110     0.926 
sex             0.975     0.283     3.444     4.250     0.024    32.504     0.978

                            Estimate 
Intercept~~Intercept|school    0.351 
Residual~~Residual             1.153 
ICC|school                     0.233

@@

             Estimate Std.Error   t.value        df   P(>|t|)       RIV       FMI 
(Intercept)     5.011     0.410    12.222     4.226     0.000    35.955     0.980 
sex             0.928     0.381     2.434     4.168     0.069    48.221     0.985

                            Estimate 
Intercept~~Intercept|school    0.313 
Residual~~Residual             1.428 
ICC|school                     0.188

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- glmer(mod, family = "binomial", data = na.omit(dat)) # fit the model
R> tidy(fit, conf.int = TRUE, exponentiate = TRUE)             # print estimates
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 11 x 9
   effect   group term        estim~1 std.er~2 stati~3   p.value conf.~4 conf.~5
   <chr>    <chr> <chr>         <dbl>    <dbl>   <dbl>     <dbl>   <dbl>   <dbl>
 1 fixed    <NA>  (Intercept)  0.0863  0.0182   -11.6   2.98e-31  0.0571   0.130
 2 fixed    <NA>  typeRCT      0.757   0.137     -1.54  1.22e- 1  0.531    1.08 
 3 fixed    <NA>  age          1.03    0.00265   12.9   7.39e-38  1.03     1.04 
 4 fixed    <NA>  motor_scor~  0.651   0.0732    -3.82  1.34e- 4  0.522    0.811
 5 fixed    <NA>  motor_scor~  0.489   0.0555    -6.30  2.97e-10  0.391    0.611
 6 fixed    <NA>  motor_scor~  0.274   0.0321   -11.0   2.28e-28  0.218    0.345
 7 fixed    <NA>  pupilNone    3.20    0.317     11.7   8.19e-32  2.63     3.88 
 8 fixed    <NA>  pupilOne     1.75    0.195      5.06  4.27e- 7  1.41     2.18 
 9 fixed    <NA>  ctIII        2.41    0.268      7.89  3.05e-15  1.94     2.99 
10 fixed    <NA>  ctIV/V       2.30    0.214      8.95  3.55e-19  1.92     2.76 
11 ran_pars name  sd__(Inter~  0.230  NA         NA    NA        NA       NA    
# ... with abbreviated variable names 1: estimate, 2: std.error, 3: statistic,
#   4: conf.low, 5: conf.high
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- imp %>% 
+   with(glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial")) 
R> tidy(pool(fit))
\end{CodeInput}
\begin{CodeOutput}
                     term    estimate   std.error  statistic      p.value
1             (Intercept) -2.35201485 0.340106796  -6.915519 4.943219e-12
2                    type -0.41266362 0.180240358  -2.289518 2.206828e-02
3                     age  0.03049017 0.001570142  19.418732 1.231236e-81
4 as.factor(motor_score)2 -0.66765094 0.068736642  -9.713174 3.473396e-22
5 as.factor(motor_score)3 -1.05520421 0.070218236 -15.027495 2.532464e-50
6 as.factor(motor_score)4 -1.51238926 0.072304429 -20.916966 1.894925e-90
7                   pupil  0.48421322 0.038983420  12.421004 6.807219e-17
8                      ct  0.43474506 0.029967898  14.507025 2.318762e-36
             b          df dfcom         fmi      lambda m         riv
1 5.283158e-04 10118.53147 11013 0.005677319 0.005480804 5 0.005511008
2 4.875905e-05 10894.06456 11013 0.001984282 0.001801077 5 0.001804327
3 3.333802e-08  6323.24413 11013 0.016538196 0.016227183 5 0.016494848
4 4.185462e-05  8330.19444 11013 0.010867814 0.010630361 5 0.010744580
5 5.135737e-05  7632.09368 11013 0.012757930 0.012499256 5 0.012657465
6 1.404727e-04  2826.80320 11013 0.032927570 0.032243595 5 0.033317884
7 3.572093e-04    49.95954 11013 0.309174397 0.282061687 5 0.392877330
8 8.584367e-05   294.82759 11013 0.120648604 0.114703577 5 0.129565165
          ubar
1 1.150387e-01
2 3.242808e-02
3 2.425341e-06
4 4.674500e-03
5 4.868972e-03
6 5.059363e-03
7 1.091056e-03
8 7.950625e-04
\end{CodeOutput}
\begin{CodeInput}
R> as.mitml.result(fit)
\end{CodeInput}
\begin{CodeOutput}
[[1]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
      AIC       BIC    logLik  deviance  df.resid 
10495.423 10561.192 -5238.712 10477.423     11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2843  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.37193                 -0.41015                  0.03052  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.65803                 -1.04612                 -1.51245  
                  pupil                       ct  
                0.50405                  0.42496  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[2]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
     AIC      BIC   logLik deviance df.resid 
10500.88 10566.65 -5241.44 10482.88    11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2917  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.37717                 -0.41511                  0.03067  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.66935                 -1.05210                 -1.49428  
                  pupil                       ct  
                0.49013                  0.43835  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[3]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
      AIC       BIC    logLik  deviance  df.resid 
10505.026 10570.795 -5243.513 10487.026     11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2908  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.32339                 -0.42359                  0.03023  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.67142                 -1.05776                 -1.51038  
                  pupil                       ct  
                0.49756                  0.42474  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[4]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
      AIC       BIC    logLik  deviance  df.resid 
10519.511 10585.280 -5250.755 10501.511     11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2961  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.33576                 -0.40872                  0.03039  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.66477                 -1.05453                 -1.51860  
                  pupil                       ct  
                0.45928                  0.44419  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 1 lme4 warnings

[[5]]
Generalized linear mixed model fit by maximum likelihood (Laplace
  Approximation) [glmerMod]
 Family: binomial  ( logit )
Formula: mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 |  
    name)
      AIC       BIC    logLik  deviance  df.resid 
10522.038 10587.807 -5252.019 10504.038     11013 
Random effects:
 Groups Name        Std.Dev.
 name   (Intercept) 0.2955  
Number of obs: 11022, groups:  name, 15
Fixed Effects:
            (Intercept)                     type                      age  
               -2.35182                 -0.40575                  0.03064  
as.factor(motor_score)2  as.factor(motor_score)3  as.factor(motor_score)4  
               -0.67468                 -1.06551                 -1.52623  
                  pupil                       ct  
                0.47006                  0.44148  
optimizer (Nelder_Mead) convergence code: 0 (OK) ; 0 optimizer warnings; 2 lme4 warnings

@@

@@

\begin{itemize}
\item
  Adding evaluations of the imputations such as convergence checks
\item
  Adding some kind of help function to mice that suggests a suitable
  predictor matrix to the user, given a certain analysis model.
\item
  Adding a \texttt{multilevel\_ampute()} wrapper function in mice.
\item
  Exporting \texttt{mids} objects to other packages like \texttt{lme4}
  or \texttt{coxme}?
\item
  Adding a ICC=0 dataset to show that even if there is no clustering it
  doesn't hurt.
\item
  Show use case for deductive imputation for cluster level variables?
\item
  env dump in repo
\item
  I don't know if in your article you cover something about model
  complexity, for example sometimes I have to switch from 2l. methods to
  1l. methods just because the model didn't converge.. this is due to
  the considered imputation model is very complex regarding the amount
  of information counted\ldots{} I know that a solution for an
  imputation model with many predictors is to check correlation plots as
  you did or use quickpred().. (maybe you can add this somewhere after
  the correlation plots)\ldots..But as for cluster specification, I
  don't know if besides plots of distribution per cluster there is
  something else can be done to see if i have to use 1l. or 2l. for a
  given variable, also I have no idea.. how to test which is better
  between 2l.norm or 2l.2stage.norm.
\end{itemize}

Translated with www.DeepL.com/Translator (free version)
